## **BIOE215 Fall 2023**

## Aspen Ellis

## *Computational Thinking 1 Assessment*

(Team Porifera because I am but a babe)


**Q1**

Match the following types of real-world data to their equivalent R type and explain your reasoning.

-   Real world data: scientific names, the count of plants in a quadrat, whether or not it rained on a sequence of days, and the reaction times of birds to disturbances in seconds

    -   *scientific names = character, because it's text*

    -   *\# plants in a quadrat = integer, because it's whole numbers*

    -   *whether or not it rained = logical, because it's y/n data*

    -   *reaction times = numeric, because it's not necessarily whole numbers, could need decimals*

**Q2**

```{r}
bird_mass_g <- c(100.1, 99.2, 99.3, NA, 100.0, 101.5, 94.7, 99.2, 108.2)
mean_mass <- mean(bird_mass_g)
sd_mass <- sd(bird_mass_g)
is_outlier <- bird_mass_g > mean_mass + 3 * sd_mass
num_outliers <- sum(is_outlier)
```

-   In plain english, what does this code chunk do?

    -   *It:*

        -   *makes a vector of bird masses*

        -   *calculates the mean mass of the birds in the vector and stores it to `mean_mass`*

        -   *calculates the standard deviation of the bird mass and stores it to `sd_mass`*

        -   *defines outliers as values where the mass is greater that the mean +3\*SD*

        -   *counts the number of outliers and stores it to `num_outliers`*

-   Two of the lines have mistakes that keep the code from doing what it's supposed to. What are the mistakes and how would you fix them?

    -   *the mistakes are that the `mean_mass` and `sd_mass` lines forget to remove the NAs, so don't work. You could fix by creating a new vector with the NAs removed:*

        ```{r}
        bird_mass_g <- c(100.1, 99.2, 99.3, NA, 100.0, 101.5, 94.7, 99.2, 108.2)
        bird_mass_g <- bird_mass_g[!is.na(bird_mass_g)]
        mean_mass <- mean(bird_mass_g)
        sd_mass <- sd(bird_mass_g)
        is_outlier <- bird_mass_g > mean_mass + 3 * sd_mass
        num_outliers <- sum(is_outlier)
        ```

-   What are the types of `bird_mass_g`, `is_outlier`, and `num_outliers`?

    -   *`bird_mass_g` is numeric*

    -   *`is_outlier` is logical*

    -   *`num_outliers` is integer*

**Q3**

How could you change the first line of code in the following chunk so that `median_count` comes out to `5L`?

```{r}
quad_counts <- c(2L, 19L, 4L, 5L, 5L, 12L, 5L, 7L)
valid_quads <- c(1, 2, 3, 5, 7) 
quad_counts2 <- quad_counts[valid_quads]
median_count <- median(quad_counts2)
```

I just changed the 0L to a 5L so that the median position was a 5L when I sampled the 1st, 2nd, 3rd, 5th, and 7th positions
